canaccessR: An open data product for analyzing transportation accessibility to employment and grocery stores in Canada's largest metropolitan areas.

### 2025-03-11 paper/Manuscript-Data-Package.Rmd
finish descriptive and start plot
finish descriptive statistics table and discussion; started creating dataset for plotting on the how to use session

```{r load-packages, include=FALSE, cache=FALSE}
# Load packages

```{r load-data, include = FALSE}

@@

Below, we present some descriptive statistics from the travel time matrices contained in the {canaccessR} package.

```{r}

@@

This section presents some potential applications of the data package.

>>>>

```{r load-packages, include=FALSE, cache=FALSE}
# Load packages
library(tidyverse)
library(data.table)
# library(summarytools)
library(gt)
library(gtsummary)
# library(sf)
# library(tmap)
# library(tmaptools)
```

```{r load-data, include = FALSE}
load("sk_full.Rdata")
```

@@

We now present some descriptive statistics from the travel time matrices contained in the {canaccessR} package. 
The Table @tbl-table_1 below summarizes the travel time estimates.
The travel time datasets contain, in total, 97,784,850 origin-destination pairs (observations) from population to employment locations for all the DA's in the sample and 18,519,897 pairs from population to groceries stores.
Considering all areas combined, the travel times estimated to jobs have an average of 78, standard deviation of 25, 1st quantile of 59, median of 80, and 3rd quantile of 99.
For travel times to groceries stores, these same statistics are the 76, 26, 56, 76, and 97, respectively.

```{r}
#| label: tbl-table_1
#| echo: false
#| message: false

sk_full %>% 
  gt() %>% 
  fmt_number(use_seps = TRUE, decimals = FALSE) |>  
  tab_options( 
    table.font.size = 10,
    table.width = pct(100) 
    )

@@

This section exemplifies how the package can be used through visual representation.

A more thorough example of the package's use can be found on the School of Cities' recent report on Canada's Urban Infrastructure Deficit.
In its 11th Chapter, we use the travel time matrices to estimate accessibility metrics to jobs and groceries stores before and after the pandemic [@pargaDemocraticAccessOur2024].
We then compare how changes affected groups differently according to their spatial distribution and income level, thus making explicit the connection of the package's information and matters of equity in transportation.
The report is freely available for download at the State of Cities Summit [website](https://stateofcitiessummit.ca/report).

### 2025-03-11 paper/Manuscript-Data-Package.tex
finish descriptive and start plot
finish descriptive statistics table and discussion; started creating dataset for plotting on the how to use session

@@

Below, we present some descriptive statistics from the travel time
matrices contained in the \{canaccessR\} package.

>>>>

\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\usepackage{colortbl}
\usepackage{array}
\usepackage{anyfontsize}
\usepackage{multirow}

@@

We now present some descriptive statistics from the travel time matrices
contained in the \{canaccessR\} package. The Table \citet{tbl-table_1}
below summarizes the travel time estimates. The travel time datasets
contain, in total, 97,784,850 origin-destination pairs (observations)
from population to employment locations for all the DA's in the sample
and 18,519,897 pairs from population to groceries stores. Considering
all areas combined, the travel times estimated to jobs have an average
of 78, standard deviation of 25, 1st quantile of 59, median of 80, and
3rd quantile of 99. For travel times to groceries stores, these same
statistics are the 76, 26, 56, 76, and 97, respectively.

\begingroup
\setlength\LTleft{0\linewidth}
\setlength\LTright{0\linewidth}\fontsize{7.5pt}{9.0pt}\selectfont
\begin{longtable}{@{\extracolsep{\fill}}clrrrrrr}
\toprule
Study Region Name & Destination & Observations & Mean & Sd & P25 & P50 & P75 \\ 
\midrule\addlinespace[2.5pt]
All regions & Employment & 97,784,850 & 78 & 25 & 59 & 80 & 99 \\ 
Toronto & Employment & 41,038,062 & 82 & 25 & 64 & 85 & 103 \\ 
Montréal & Employment & 30,091,411 & 76 & 25 & 57 & 77 & 97 \\ 
Vancouver & Employment & 12,254,478 & 76 & 25 & 58 & 77 & 96 \\ 
Calgary & Employment & 3,022,244 & 75 & 23 & 59 & 75 & 91 \\ 
Ottawa & Employment & 2,936,912 & 74 & 24 & 56 & 74 & 93 \\ 
Edmonton & Employment & 2,057,486 & 70 & 23 & 53 & 70 & 87 \\ 
Québec City & Employment & 1,473,702 & 72 & 25 & 54 & 72 & 93 \\ 
Winnipeg & Employment & 1,574,205 & 62 & 22 & 46 & 61 & 76 \\ 
Hamilton & Employment & 2,089,172 & 82 & 28 & 61 & 87 & 106 \\ 
Waterloo & Employment & 552,826 & 71 & 27 & 49 & 69 & 93 \\ 
London & Employment & 415,174 & 61 & 21 & 46 & 60 & 74 \\ 
Halifax & Employment & 279,178 & 66 & 25 & 47 & 65 & 84 \\ 
All regions & Groceries Stores & 18,519,897 & 76 & 26 & 56 & 76 & 97 \\ 
Toronto & Groceries Stores & 8,512,874 & 80 & 25 & 61 & 82 & 101 \\ 
Montréal & Groceries Stores & 2,993,965 & 73 & 26 & 53 & 74 & 94 \\ 
Vancouver & Groceries Stores & 4,540,106 & 73 & 25 & 53 & 74 & 93 \\ 
Calgary & Groceries Stores & 502,757 & 70 & 22 & 54 & 70 & 86 \\ 
Ottawa & Groceries Stores & 619,791 & 72 & 24 & 55 & 72 & 91 \\ 
Edmonton & Groceries Stores & 324,030 & 68 & 23 & 51 & 67 & 84 \\ 
Québec City & Groceries Stores & 234,600 & 70 & 26 & 51 & 70 & 89 \\ 
Winnipeg & Groceries Stores & 362,566 & 58 & 21 & 43 & 57 & 72 \\ 
Hamilton & Groceries Stores & 281,118 & 84 & 29 & 60 & 91 & 110 \\ 
Waterloo & Groceries Stores & 44,726 & 66 & 28 & 44 & 62 & 90 \\ 
London & Groceries Stores & 52,617 & 55 & 19 & 42 & 55 & 67 \\ 
Halifax & Groceries Stores & 50,747 & 63 & 26 & 42 & 63 & 83 \\ 
\bottomrule
\end{longtable}
\endgroup
